install.packages("sf", dependencies = TRUE) install.packages("terra", dependencies = TRUE) install.packages(c("spData", "mapview", "ggplot2", "tmap")) install.packages("tidyverse", dependencies = TRUE) r

library(sf) library(spData)

data("cycle_hire_osm", package = "spData") head(cycle_hire_osm) \# plot first few rows

plot(cycle_hire_osm["geometry"], axes = TRUE) st_crs(cycle_hire_osm) cycle_hire_osm \<- ßst_transform(cycle_hire_osm, crs = st_crs(32630)) \# UTM zone 30N for London head(cycle_hire_osm) library(mapview) mapview(cycle_hire_osm["capacity"], layer.name = "Rental capacity")

```{r}
library(sf)
data("seine", package = "spData")
seine
```

```{r}
library(ggplot2)

ggplot(seine) +
  geom_sf(aes(color = name)) +
```

```{r}
geom_sf_label(aes(label = name)) # add labels for river names
```

```{r}
data(world, package = "spData")
head(world)
breaks <- quantile(world$gdpPercap, probs = c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm = TRUE) # get quantile values

ggplot(world) + geom_sf(aes(fill = gdpPercap)) + 
  scale_fill_gradientn(colours= c("#ffffcc", "#a1dab4" , "#41b6c4", "#2c7fb8", "#253494"), # binned color scale
                       trans = 'log', # log-transformation
                       breaks = breaks) +
  theme_minimal()
```

Geometry operations

```{r}
#Locate the center point of geometries
world_pts <- st_centroid(world)

```

```{r}
#Simplify shapes by removing vertices
world_projected <- st_transform(world, "EPSG: 3857") # need to transform to projected CRS
world_simple <- st_simplify(world_projected, dTolerance = 1e5) # 100km simplification
```

```{r}
#Cast the simplified POLYGON objects back to type 'MULTIPOLYGON', and remove the rows with empty geometries

world_simple <- st_cast(world_simple, "MULTIPOLYGON")
world_simple <- world_simple[!st_is_empty(world_simple),] # subset rows

```

```{r}
#Extend the boundaries of vectors
world_buffered <- st_buffer(world_simple, dist = 1e5) # 100km
```

```{r}
#Remove overlaps among the polygons in world_simple and merge all polygons
world_diff <- st_difference(world_simple) # exclude overlaps
world_union <- st_union(world_simple)
```

```{r}
#Plot the manipulated objects derived from the country polygons
#packages
library(sf)
library(tmap)
sf_use_s2(FALSE)

tm_shape(world_pts) + tm_dots() + # polygon centroids
  tm_shape(world_simple) + tm_polygons(alpha = 0.5) + # simplified
  tm_shape(world_buffered) + tm_polygons(alpha = 0.5) + # buffered
  tm_shape(world_union) + tm_polygons(alpha = 0.5) + # merged (union)
  tm_view(set.view = 2) # zoom in

```
------------------------------------------------------------------------

#learning  about geometry operations 

#Locating the center point of geometries
```{r}
world_pts <- st_centroid(world)

```
```{r}
library(sf)

# Cast the simplified POLYGON objects back to MULTIPOLYGON
world_simple <- st_cast(world_simple, "MULTIPOLYGON")
world_simple <- world_simple[!st_is_empty(world_simple), ]  # subset rows

# Simplify shapes by removing vertices
world_projected <- st_transform(world, "EPSG:3857")         # project to CRS
world_simple    <- st_simplify(world_projected, dTolerance = 1e5)


```
```{r}
#Extend the boundaries of vectors
world_buffered <- st_buffer(world_simple, dist = 1e5) # 100km
```

```{r}
#Remove overlaps among the polygons in `world_simple` and merge all polygons
world_diff <- st_difference(world_simple) # exclude overlaps
world_union <- st_union(world_simple)
```

------------------------------------------------------------------------

#Plot the manipulated objects derived from the country polygons

```{r}
library(sf)
library(tmap)

# (optional) if you want planar ops
sf_use_s2(FALSE)

# --- data prep ---
# project to Web Mercator
world_projected <- st_transform(world, 3857)

# simplify (keep topology) and fix invalid geometries
world_simple <- st_simplify(world_projected, dTolerance = 1e5, preserveTopology = TRUE)
world_simple <- st_make_valid(world_simple)

# derived layers
world_buffered <- st_buffer(world_simple, dist = 1e5)   # ~100 km buffer
world_union    <- st_union(world_simple)                # merged polygons
world_pts      <- st_centroid(world_simple)             # polygon centroids

# --- map ---
tmap_mode("view")
tm_shape(world_pts)      + tm_dots() +
tm_shape(world_simple)   + tm_polygons(alpha = 0.5) +
tm_shape(world_buffered) + tm_polygons(alpha = 0.5) +
tm_shape(world_union)    + tm_polygons(alpha = 0.5) +
tm_view(set.view = 2)    # zoom in
```


# Subset rows of an `sf` ('target') object based on its spatial relationship with another

```{r message=FALSE, error=FALSE, warning = FALSE, fig.height=1, out.width = "100%", dpi=300, fig.align='center'}

uk <- st_filter(world_projected, # target: world polygons
                st_transform(cycle_hire_osm, st_crs(world_projected)), # bicycle rental points in London - crs must be the same
                .predicate = st_intersects) # default setting

tm_shape(uk) + tm_polygons() +
  tm_shape(cycle_hire_osm) + tm_dots()

```

# Match rows (and append columns) to an `sf` ('target') object based on its spatial relationship with another

```{r message=FALSE, error=FALSE, warning = FALSE, fig.height=1, out.width = "100%", dpi=300, fig.align='center'}
cycle_hire_osm <- st_join(cycle_hire_osm, # target
                          st_transform(world_projected, st_crs(cycle_hire_osm)))

head(cycle_hire_osm)
```

#Combine the official dataset `cycle_hire` with the OSM dataset `cycle_hire_osm` of bicycle rental points across London, but only those beyond a distance of 10 m (remove duplicate points)

```{r echo = FALSE, message=FALSE, error=FALSE, warning = FALSE, fig.height=1, out.width = "100%", dpi=300, fig.align='center'}

library(dplyr)

data("cycle_hire", package = "spData")
cycle_hire <- st_transform(cycle_hire, st_crs(cycle_hire_osm))

predicate <- st_is_within_distance(cycle_hire_osm, 
                                   cycle_hire, 
                                   dist = 10)

to_add <- !lengths(predicate) > 0

cycle_hire_osm_subset <- cycle_hire_osm[to_add,] %>% 
  mutate(source = "osm")

cycle_combined <- cycle_hire %>% 
  mutate(source = "official") %>% 
  bind_rows(cycle_hire_osm_subset)

tm_shape(cycle_combined) + tm_dots("source")

```

------------------------------------------------------------------------

# Working with rasters {.tabset .tabset-fade .tabset-pills}

#Load package

```{r message=FALSE, error=FALSE}
library(terra)
```

## Continuous

#Example raster: Elevation in Luxembourg

```{r message=FALSE, error=FALSE, fig.width=6, fig.height=4.5, out.height=280, dpi=300, fig.align='center'}
filepath <- system.file("ex/elev.tif", package = "terra")
elevation <- rast(filepath)
elevation_utm <- project(elevation, "EPSG:23032") # project to UTM zone 32N for Luxembourg

```

#Example vector: Polygons of Luxembourg

```{r message=FALSE, error=FALSE, fig.width=8, fig.height=4.5, out.height=300, dpi=300, fig.align='center'}
filepath <- system.file("ex/lux.shp", package = "terra")
luxembourg <- vect(filepath)
luxembourg
```

#Summarise the mean raster values for each sub-region (polygon) in Luxembourg

```{r}
elev_mean <- extract(elevation, luxembourg, 
                     fun = mean, na.rm = TRUE)
luxembourg$elevation_mean <- elev_mean$elevation # add results to new column
```

#Plot the rasters and polygons

```{r message=FALSE, error=FALSE, fig.height=1, out.width = "100%", dpi=300, fig.align='center'}

tm_shape(elevation) +
  tm_raster(palette = "RdYlGn", n = 10) + # use similar color palette
  tm_shape(st_as_sf(luxembourg)) + # convert to sf, as tmap currently does not support SpatVect
  tm_polygons(col = "elevation_mean", palette = "RdYlGn", alpha = 0.7) 
```

#Narrow down raster data to a specific area of interest

```{r message=FALSE, error=FALSE, fig.height=5, out.width="70%", dpi=300, fig.align='center'}
luxembourg_subset <- luxembourg[luxembourg$NAME_2 == "Mersch"]
elevation_subset <- crop(elevation, luxembourg_subset) # reduce spatial extent
elevation_subset <- mask(elevation_subset, luxembourg_subset) # convert values outside polygon to NA
plot(elevation_subset)
```

------------------------------------------------------------------------

## Discrete

#Classify values into a discrete (categorical) raster

```{r message=FALSE, error=FALSE, fig.height=5, out.width="70%", dpi=300, fig.align='center'}
elevation_2class <- ifel(elevation > 400, 1, 0) # E.g., threshold for 'high' elevation at 400m
plot(elevation_2class) # low/high
```

```{r message=FALSE, error=FALSE, fig.height=5, out.width="70%",  dpi=300, fig.align='center'}
elevation_3class <- 
  classify(elevation, 
           rbind(c(400, Inf, 2), # matrix of 3 columns, 'from', 'to' and 'becomes'
                 c(300, 400, 1),
                 c(-Inf, 300, 0)))  
plot(elevation_3class) # low/mid/high
```

------------------------------------------------------------------------

# Data conversions

#Vectorize rasters into points, lines and polygons

```{r message=FALSE, error=FALSE}
elevation_pts <- as.points(elevation_subset) # points
elevation_contours <- as.contour(elevation_subset) # lines
elevation_polygons <- as.polygons(elevation_subset) # polygons
```

#Visualise all output

```{r message=FALSE, error=FALSE, fig.height=1, out.width="100%", dpi=300, fig.align='center'}

tm_shape(elevation_subset) + tm_raster(palette = "RdYlGn", n = 10) + # original raster
  tm_shape(st_as_sf(elevation_pts)) + tm_dots() + 
  tm_shape(st_as_sf(elevation_contours)) + tm_lines() + 
  tm_shape(st_as_sf(elevation_polygons)) + tm_polygons(alpha = 0) + 
  tm_view(set.view = 11)
```

#Rasterize the points dataset `cycle_hire_osm` based on the column '`capacity`'

```{r message=FALSE, error=FALSE}
# first create a raster template at a specified resolution and crs
rast_template <- rast(ext(cycle_hire_osm), 
                      resolution = 1000,
                      crs = st_crs(cycle_hire_osm)$wkt) # input format is WKT

# then, rasterize the points dataset based on the template
cyclehire_capacity <- 
  rasterize(cycle_hire_osm, rast_template,
            field = "capacity", 
            fun = sum, # add up values of points within each pixel
            na.rm = TRUE)
```

# Overlay the original points over the per-pixel capacity of bicycle docking stations

```{r message=FALSE, error=FALSE, fig.height=1, out.width="100%", dpi=300, fig.align='center'}

tm_shape(cyclehire_capacity) + 
  tm_raster(title = "Capacity (sum)", palette = "YlGn") +
  tm_shape(st_as_sf(cycle_hire_osm)) + 
  tm_dots(size = 0.01) 
```



